---
title: "solar_model"
author: "sean connin"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE) # set global options

```

Install libraries

```{r}
library(tidyverse)
library(RSocrata) # streamline api request
library(magrittr)
library(janitor)
library(usethis) # set up Renviron for project
library(dlookr)
library(flextable)
library(lubridate)
```
Set up credentials - api token

* edit_r_environ()  #open Renviron and store api token as 

```{r}

api<-Sys.getenv("api_token")

```

Acquire data via. api

Note: beginning dimensions (120247, 32)

```{r}

# query endpoint and filter based on program status and type

url <- "https://data.ny.gov/resource/3x8r-34rs.json?Project Status=Complete&Program Type=Residential/Small Commercial"

df<-read.socrata(url, app_token = api)  # credential stored in Renviron

closeAllConnections()

```
clean and subset data 

Note that georeference.coordinates is a vector of lists. each containing lat and long. 

We can unnest if necessary:

* unnest_wider(georeference.coordinates)%>%  # use rectangle package in purrr to unnest list elements
    
* rename(longitude = ...1, latitude = ...2)%>%   # assign names to unnested cols

We will remove from working dataset and can add back in later if required

```{r}


df_mod<- tibble(df)%>%
    
    drop_na(project_number, date_application_received, date_install, project_cost, project_status)%>%
    
    select(!c(reporting_period, state, program_type, project_status, legacy_project_number, primary_inverter_model_number, pv_module_model_number, affordable_solar, green_jobs_green_new_york_participant, georeference.type, georeference.coordinates))%>%
    
    janitor::remove_empty(which = c("rows", "cols"))%>% # remove any rows/cols with no values 
    
    distinct()%>%     # remove duplicate rows

    mutate_if(is_character, str_replace_all, '\\$|,|z_|<', '')%>%   # clean any special chars in character variables
    
    relocate(county, .after = city)%>%
    
    mutate_if(is_character, str_replace_all, "No|no",'N')%>%
    
    mutate_if(is_character, str_replace_all, "Yes|yes",'Y')%>%
    
    mutate_if(is_character, str_replace_all, "Nn-Residential",'Commercial')%>%
    
    mutate_at(vars(inverter_quantity, pv_module_quantity, expected_kwh_annual_production), funs(as.integer))%>% 
    
    mutate_at(vars(project_cost, totalnameplatekwdc, incentive), funs(as.double))%>%
    
    mutate_at(vars(electric_utility, sector, purchase_type), funs(as_factor))  #as_factor is forecats


```

# evaluate count and proportion of missingness

Notes:

* We are missing 64% of the col 'remote_net_metering'; we will remove this column
* Given the large obs count = 120247 and low level of missingness (< 5%) within cols, we will drop rows with missing entries

```{r}

# create table to describe missing data

df_mod%>%
    diagnose()%>%
    dplyr::select(-unique_count, -unique_rate)%>%
    filter(missing_count>0)%>%
    arrange(desc(missing_count))%>%
    flextable()

# drop remote_net_metering and rows with NA

df_mod%<>%
    dplyr::select(!c(remote_net_metering))%>%
    drop_na()

# confirm removal of all rows/cols with NA

sum(is.na(df_mod))

```
Engineer project_days as date_install - date_application_received using lubridate package

```{r}
df_mod%<>%
    
    mutate(year_completed = year(date_install), .after=date_install)%>%  
    
    mutate(days=difftime(date_install, date_application_received, units = "days"), .after=year_completed)%>%
    
    separate(days, into = 'completion_days', sep = " ")%>%
    
    mutate_at(vars(completion_days), funs(as.integer))

df_mod%<>%
    select(!c(date_application_received, date_install))

View(df_mod)

```

Partition dataset into residential and commercial tibbles

```{r}
df_mod%<>%
```


Run additional filters for data cleaning

```{r}

# subset residential sector and eval numeric types

df_mod%>%
    filter(sector %in% 'Residential')%>%
    diagnose_numeric()%>%
    dplyr::select(variables, min, mean, median, max, zero, minus)%>%
    flextable(theme_fun = theme_booktabs())

# subset commercial sector and eval numeric types

df_mod%>%
    filter(sector %in% 'Commercial')%>%
    diagnose_numeric()%>%
    dplyr::select(variables, min, mean, median, max, zero, minus)%>%
    flextable(theme_fun = theme_booktabs())

# subset residential sector and eval categorical types

df_mod%>%
    filter(sector %in% 'Residential')%>%
    diagnose_category()%>%
    flextable(theme_fun = theme_booktabs())

# subset commerical sector and eval categorical types

df_mod%>%
    filter(sector %in% 'Commercial')%>%
    diagnose_category()%>%
    flextable(theme_fun = theme_booktabs())

# remove rows where completion_days = 0 or less, based on diagnose_numeric results

df_mod%<>%
    filter(completion_days > 0 & project_cost > 0 & inverter_quantity >0 & totalnameplatekwdc > 0)








plot(df_mod$completion_days)


df_mod%>%find_skewness(index=FALSE, thres=TRUE)

df_mod%>%normality() # wilkes-shapiro

df_mod%>%plot_normality()

unique(df_mod$sector)
```
outliers

```{r}

df_mod%>%
    filter(sector %in% 'Residential')%>%
    dplyr::select(find_outliers(df_mod, index = FALSE)) %>% 
    plot_outlier()



df_mod%>%
    filter(sector %in% 'Commercial')%>%
    dplyr::select(find_outliers(df_mod, index = FALSE)) %>% 
    plot_outlier()
    

df_mod%>%
    filter(sector %in% 'Residential')%>%
    diagnose_outlier()%>%
    flextable() 

df_mod%>%
    filter(sector %in% 'Commercial')%>%
    diagnose_outlier()%>% 
    flextable() 


```

